### 2022-09-19 docs/paper.tex
Add paper template

\documentclass[a4paper]{article}

\author{Jelmer Firet, Bram Pulles}
\title{\textbf{Automated Reasoning\\Practical Assignment -- Part 1}}

\begin{document}
\maketitle

\section{Pallets}

\section{Chip design}

\section{Dinner}

\section{Program safety}

\end{document}